Finiteness in Cubical Type Theory

### 2020-06-17 sections/introduction.tex
better spacing

The are two broad ways to define types in Agda: as an inductive
\(\AgdaKeyword{data}\) type, similar to data type definitions in Haskell, or as
a \(\AgdaKeyword{record}\).
Here we'll define the basic type formers used in MLTT.\@
\begin{definition}[Basic Types]
  The three basic types---often called 0, 1, and 2 in MLTT---here will be
  denoted with their more common names: \(\bot\), \(\top\), and
  \(\mathbf{Bool}\), respectively.
  \begin{multicols}{3} \centering
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{bot}
    \end{agdalisting}

@@

  \begin{center}
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma}
    \end{agdalisting}
  \end{center}
  We will use different notations to refer to this type depending on the
  setting.
  The following four expressions all denote the same type:
  \begin{multicols}{4}
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-1}
    \end{agdalisting}

    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-3}
    \end{agdalisting}

    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-4}
    \end{agdalisting}

    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-2}
    \end{agdalisting}
  \end{multicols}
  The non-dependent product is a special instance of the dependent.
  We denote a simple pair of types \(A\) and \(B\) as \(A \times B\).
\end{definition}
\begin{definition}[Dependent Product]
  Dependent products (dependent functions) use the \(\Pi\) symbol.
  The three following expressions all denote the same type:
  \begin{multicols}{3}
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{pi-syntax-1}
    \end{agdalisting}

    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{pi-syntax-2}
    \end{agdalisting}

    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{pi-syntax-3}
    \end{agdalisting}
  \end{multicols}
  Non-dependent functions are denoted with the arrow (\(\rightarrow\)).
\end{definition}

At this point, as a quick example, we can define the first of our objects for
the countdown transformation: the vector of Booleans for selection.
A vector is relatively simple to define: a vector of zero elements is simply a
unit, a vector of \(n+1\) elements is the product of an element and a vector of
\(n\) elements.
\begin{agdalisting}
  \ExecuteMetaData[agda/Data/Vec/Iterated.tex]{vec-def}
\end{agdalisting}
From this we can see that a vector of \(n\) Booleans has the type
\(\AgdaDatatype{Vec} \; \AgdaDatatype{Bool} \; n\).

>>>>

The are two broad ways to define types in Agda: as an inductive
\(\AgdaKeyword{data}\) type, similar to data type definitions in Haskell, or as
a \(\AgdaKeyword{record}\).
Here we'll define the basic type formers used in MLTT.\@
\begin{definition}[Basic Types]
  The three basic types---often called 0, 1, and 2 in MLTT---here will be
  denoted with their more common names: \(\bot\), \(\top\), and
  \(\mathbf{Bool}\), respectively.
  \begin{multicols}{3}
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{bot}
    \end{agdalisting}

@@

  \begin{center}
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma}
    \end{agdalisting}
  \end{center}
  We will use different notations to refer to this type depending on the
  setting.
  The following four expressions all denote the same type:
  \begin{multicols}{4}
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-1}
    \end{agdalisting} \columnbreak
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-3}
    \end{agdalisting} \columnbreak
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-4}
    \end{agdalisting} \columnbreak
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-syntax-2}
    \end{agdalisting}
  \end{multicols}\vspace{-1.5\baselineskip}%
  The non-dependent product is a special instance of the dependent.
  We denote a simple pair of types \(A\) and \(B\) as \(A \times B\).
\end{definition}
\begin{definition}[Dependent Product]
  Dependent products (dependent functions) use the \(\Pi\) symbol.
  The three following expressions all denote the same type:
  \begin{multicols}{3}
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{pi-syntax-1}
    \end{agdalisting} \columnbreak
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{pi-syntax-2}
    \end{agdalisting} \columnbreak
    \begin{agdalisting}
      \ExecuteMetaData[agda/Snippets/Introduction.tex]{pi-syntax-3}
    \end{agdalisting}
  \end{multicols}\vspace{-1.5\baselineskip}
  Non-dependent functions are denoted with the arrow (\(\rightarrow\)).
\end{definition}

At this point, as a quick example, we can define the first of our objects for
the countdown transformation: the vector of Booleans for selection.
A vector is relatively simple to define: a vector of zero elements is simply a
unit, a vector of \(n+1\) elements is the product of an element and a vector of
\(n\) elements.
\begin{agdalisting}
  \ExecuteMetaData[agda/Data/Vec/Iterated.tex]{vec-def}
\end{agdalisting}
From this we can see that a vector of \(n\) Booleans has the type
\(\AgdaDatatype{Vec} \; \AgdaDatatype{Bool} \; n\)
